### 2023-01-05 code/raking_and_model.qmd
need to write model equation, fix model and read about random effects

---
title: "Raking and model"
format: html
editor: visual
---

This document pertains the corrections on the clean dataset created in `Data_cleaning_and_covariates_Jan_05.qmd`. There are differences in the proportions of the dataset with regard to income, age groups, race/ethnicity, and geographical location when compared to the Census data. The dataset will be corrected using _raking_ with the package `survey`.

# Raking

This section follows the steps provided by [https://sdaza.com/blog/2012/raking/](https://sdaza.com/blog/2012/raking/) https://www.r-bloggers.com/2018/12/survey-raking-an-illustration/

The variables to define raking weights are:

  - income
  - age groups 
  - geographical location
  - race and ethnicity

  The population proportions in each case have been already presented in the data cleaning document.

  The dataset has missing observations in the race (28% missing), and second_dose (26.3%) covariates. First attempt for raking and model will after removing all missing observations.

```{r,define-target-list}

library(weights)
library(survey)
library(magrittr)

clean_data<-read.csv(here("data","clean_dataset.csv"))

test<-clean_data%>%drop_na(race)

#remove municipal status as well to avoid problems

test<-test%>%
  select(-Municipal.status)

#make factors
test<-test %<>%
  mutate_at(c("income_ord","race","age_group","location","Geographic.area"),factor)

test<-test %>%
  mutate(first_dose_m=case_when(
    first_dose=="yes"~1,
    first_dose=="no"~0
  ))

#get percentages
wpct(test$income_ord)
wpct(test$age_group)
wpct(test$race)
wpct(test$location)

#create dataframes for weights

t1<-svydesign(id=~1,data=test)

ages<-data.frame(age_group=c("16_24",
                             "25_34",
                             "35_44",
                             "45_54",
                             "55_64",
                             "65_and_over"),
                 Freq=c(0.127,
                        0.129,
                        0.128,
                        0.149,
                        0.137,
                        0.167))

races<-data.frame(race=c("arab_middle_eastern",
                         "black",
                         "east_asian_pacific_islander",
                         "indigenous",
                         "latin_american",
                         "mixed",
                         "other",
                         "south_asian",
                         "white_caucasian"),
                  Freq=c(0.016,
                         0.047,
                         0.066,
                         0.028,
                         0.015,
                         0.01,
                         0.052,
                         0.087,
                         0.678
                  ))

income_ords<-data.frame(income_ord=c("under_15000",
                                     "15000_24999",
                                     "25000_39999",
                                     "40000_59999",
                                     "60000_89999",
                                     "over_90000"),
                        Freq=c(0.057,
                               0.075,
                               0.116,
                               0.154,
                               0.195,
                               0.403
                        ))

locations<-data.frame(location=c("All other cities",
                                 "Toronto"),
                      Freq=c(0.687,
                             0.313))

t.rake<-rake(t1,
             sample=list(~age_group,
                         ~race,
                         ~income_ord,
                         ~location),
             population=list(ages,races,income_ords,locations))

model<-svyglm(first_dose_m~income_ord+age_group+race,design=t1,family = quasibinomial(), control= list(maxit=50))

summary(model)

```

Test example for raking from `survey` package

```{r,test-raking}
# data(api)
# 
# dclus1 <- svydesign(id=~dnum, weights=~pw, data=apiclus1, fpc=~fpc) #no clusters
# rclus1 <- as.svrepdesign(dclus1) #create replicate weights survey design
# 
# svymean(~api00, rclus1)
# svytotal(~enroll, rclus1)
# 
# pop.types <- data.frame(stype=c("E","H","M"), Freq=c(4421,755,1018))
# pop.schwide <- data.frame(sch.wide=c("No","Yes"), Freq=c(1072,5122))
# 
# rclus1r <- rake(rclus1, list(~stype,~sch.wide), list(pop.types, pop.schwide))
# 
# svymean(~api00, rclus1r)
# svytotal(~enroll, rclus1r)
test$first_dose<-as.factor(test$first_dose)

 ggplot(clean_data %>% count(first_dose,race) %>%    # G count number in each group
         mutate(pct=n/sum(n)),  # Calculate label positions
       aes(first_dose, n,fill=race)) +
  geom_bar(stat="identity") +
   facet_wrap(~as.factor(clean_data$Geographic.area))+
  theme_classic()
```