### 2023-01-11 data_cleaning/municipalities.qmd
Added data for the health region approach

---
title: "Municipalities dataset cleaning"
format: pdf
editor: visual
---

This document is about how the municipality data was obtained and the steps used to clean it.

The dataset was obtained from the website of the province of Ontario (https://www.ontario.ca/page/list-ontario-municipalities#section-3).

The file obtained is a CSV but it needed to be cleaned to remove the unnecessary information found in the name of each city so it can be combined with the survey dataset.First, we only keep the name of the city and the municipal region they belong to, and later remove duplicate entries.

```{r,loading-first-clean}
#| echo: false
#| warning: false

library(tidyverse)
library(here)

municipalities<-read.csv(here("data","municipalities_list.csv"))

municipalities_clean <- municipalities %>%
#get only the city/county part (see https://stackoverflow.com/questions/66212344/extract-word-in-quotes-from-string)
  mutate(Full_title=str_extract(Municipality,'(?<=(?<!\\()")[^"]+(?=(?!\\))")'))

```

The second step of cleaning this dataset is the fact that it contains the names of the regions (e.g., the Regional Municipality of Peel appears listed along the geographical area of Peel, when they are the same thing). These entries need to be removed as in some cases some towns have the same name as a geographical region (e.g., Durham in Grey County, having the same name as the Regional Municipality of Durham). To do this, we will use the geographical regions listed on the [website](https://www.amo.on.ca/about-us/municipal-101/ontario-municipalities) of the Association of Municipalities of Ontario (AMO). As a clarification, because the "Districts" listed in this website do not serve any municipal government purpose, except on the case of the District Municipality of Muskoka only this last one needs to be removed from the municipality list. Lastly, on the AMO website some municipalities have their most recent status (e.g., Brant, which appears as city, but still appears listed as "County") in the website of the province of Ontario. We will load the AMO dataset (called `geographic_areas.csv`) and use to remove those entries from `municipalities_clean` that match the entries. Later, we will manually remove entries that do not have the same name but correspond to the same region.

```{r,clean-second-step}

# load AMO dataset
geographic_areas <- read.csv(here("data","geographic_areas.csv"))

#remove entries with the same name in the Ontario dataset

municipalities_clean <- anti_join(municipalities_clean,geographic_areas,by="Full_title")

# manually remove entries

exclude <-tibble(Full_title=c("Brant, County of", 
                                       "Haldimand County", 
                                       "Norfolk County",
                                       "Prince Edward, County of",
                                       "Chatham-Kent, Municipality of"))

 #remove entries

municipalities_clean <-anti_join(municipalities_clean,exclude, by="Full_title")%>%
  #remove the first column as it is no longer needed
  subset(select=-c(Municipality))

```

One additional step is to create a column that only has the city name (as the Fields Dataset only has the city name on it without the full name of "Township of", or "City of"). After that, the dataset is saved as a csv file.

```{r}

municipalities_clean <- municipalities_clean %>%
  mutate(city=sub(",.*","",Full_title))

#save the file in csv
write.csv(municipalities_clean,here("data","municipalities_clean.csv"),row.names = FALSE)

```