### 2023-02-06 manuscript/main.qmd

---
title: '**Socio-economic and geographic determinants of Covid-19 vaccination**'
author: Ariel Mundo Ortiz
format: pdf
bibliography: refs.bib
csl: jama.csl
link-citations: yes #adds links to the citations
---

# Abstract

# Keywords {-}

```{r setup, message = FALSE,include=FALSE}
#the packages needed to run the code

library(weights)
library(here)
library(survey)
library(magrittr)
library(tidyverse)
library(lme4)
library(kableExtra)
library(modelsummary)
library(gtsummary)
library(patchwork)

```

```{r} 
#| child: 01-background.qmd
```

<!-- \FloatBarrier -->

```{r} 
#| child: 02-methods.qmd
```

```{r}
#| child: 03-results.qmd
```

# References

::: {#refs}
:::