### 2023-02-08 manuscript/03-results.qmd

# Results

## Survey Results

@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields Covid-19 survey data for vaccination status and each of the covariates analyzed. The total number of entries from the survey in the dataset after cleaning was 3,551. Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it. Regarding socio-economic factors, respondents who reported living in a household with an income under CAD 15,000 or above 90,000 represented approximately 50\% of the total number of entries, with 71\% and 75\% of the respondents in each income bracket indicating to have received the first dose of the vaccine, respectively. The age group between 16-24 years had the highest representation in the survey responses (837, 23.6\% of all responses), followed by the age group of those between 65 and over (692, 19.5\% of the total); 73\% of those in the 16 to 24 years age group reported to have received the vaccine, whereas 68\% of those above 65 years indicated the same. The Health Region with the highest number of participants in the survey was Toronto, accounting for 1,342 entries (37.7\%) and a 72\% vaccination rate. Regarding race, individuals that identified as White/Caucasian represented 1313 (37\%) of all entries and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine. On the other hand, those that identified as Indigenous had the lowest vaccine uptake with 60\% indicating to have received the vaccine.

```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey"
#| message: false
#| echo: false

source(here::here("code","data_preparation.R"))

data_summary %>%
  modify_header(label = "**Variable**") %>%
  as_kable_extra(booktabs = TRUE) %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down"` argument here.
  kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)

```

## Multivariate Regression

@tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors. The reference groups in each case were set as follows: 16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income). Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.44), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.4) and Other Race/Ethnicity (a group those that identified as Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000-39,999 (OR=0.23).

Regarding the interaction effects, the interaction of the Central Health Region and income bracket between 15000 to 24999 CAD was statistically significant (OR=0.4). In age group and income, significant effects were estimated for those 65 years and older with a household income between CAD 25000-39999 (OR=7.21), and with an income between CAD 40000-59999 (OR=3.12). In the case of race and income, significant effects were estimated for individuals that identified as Arab/Middle Easterners with a family income between 15000 and 24999 (OR=5.68) or income under CAD 15000 (OR=4.21), Black individuals with a family income under CAD 15000 (OR=4.26) or between CAD 15000-24999 (OR=3.35); additionally, there were statistically significant interaction effects in the case of those that identified as belonging to a race not identified elsewhere ("Other") within multiple income brackets: under CAD 15000 (OR=3.73), between 15000 and 24999 (OR 11.2), between 25000 and 39999 CAD (OR=4.67), and with income CAD 40000-59999 (OR=9.58). Finally, there was a significant interaction effect for South Asian individuals with an income under CAD 15000 (OR=3.10).

```{r,model-table}
#| message: false
#| warning: false
#| echo: false
#| label: tbl-model
#| tbl-cap: "Multiple Regression Analysis-Predictors of Vaccination Status"

#logistic regression
source(here::here("code","regression.R"))

# create table with the model output
m1 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group ~ "Age Group")) %>%
    bold_labels()%>%
  as_kable_extra(booktabs = TRUE, longtable=T) %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down" argument here.
  kableExtra::kable_styling(latex_options = "repeat_header")
```