### 2023-02-21 data_cleaning/data_grouping_Feb_20.qmd
re-grouped data and ran model, fixed code

---
title: "**Data cleaning and model testing with grouped categories**"
subtitle: Re-working groups
author: "Ariel Mundo"
format: pdf
toc: true
toc-depth: 2
editor: source
urlcolor: blue
#bibliography: refs.bib
---

```{r, setup, echo=FALSE}
#| warning: false
library(tidyverse)
library(here)
library(lubridate)
library(gtsummary)
library(patchwork)
library(visdat)
library(survey)
library(gt)
library(scales)
library(scico)

thm1<-scale_fill_scico_d(palette="buda",begin=0.3, end=0.8, direction = -1, aesthetics = "fill")

thm2<-scale_fill_scico_d(palette="oleron",begin=0.0, end=1.0, direction = -1, aesthetics = "fill")

thm3<-scale_fill_scico_d(palette="fes",begin=0.0, end=1.0, direction = -1, aesthetics = "fill")
```

## Process

Merge income as: below 24k, 25-60k, 60k and above
Age: 16-34, 35-54, 55 and over

Interactions: Race * Income, Race * Health Region

```{r, prepare-data}

data1<-read.csv(here("data","clean_dataset.csv"))

data1 <- data1 %>% 
  drop_na(race)

data_summary2<-data1 %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income,
                        age_group,
                        race,
                        Health_Region)) %>% 
  bold_labels()

data_summary2

## Recode income

data1<-data1%>%
    mutate(income_ord = case_when(
                                  income %in% "under_15000" ~ "under_25000",
                                  income %in% "15000_24999" ~ "under_25000",
                                  income %in% "25000_39999" ~ "25000_59999",
                                  income %in% "40000_59999" ~ "25000_59999",
                                  income %in% "60000_89999" ~ "60000_and_above",
                                  income %in% "90000_109999" ~ "60000_and_above",
                                  income %in%"over_110000" ~ "60000_and_above",
                                ))%>%
  mutate(income_ord=as.factor(income_ord))

data1<-data1%>%
    mutate(age_group_ord = case_when(
                                  age_group %in% "16_24" ~ "16_34",
                                  age_group %in% "25_34" ~ "16_34",
                                  age_group %in% "35_44" ~ "35_54",
                                  age_group %in% "45_54" ~ "35_54",
                                  age_group %in% "55_64" ~ "55_and_over",
                                  age_group %in% "65_and_over" ~ "55_and_over",
                                ))%>%
  mutate(age_group_ord=as.factor(age_group_ord))

```

```{r,raking}
#remove north health regions (low numbers)

#get health regions to eliminate

data1<-data1 %>%
  select(-Municipal.status) %>%
  rename(Race=race,Age_group_ord=age_group_ord,Income=income_ord)

data1<-data1 %<>%
  mutate_at(c("Income","Race","Age_group_ord","Health_Region"),factor)

data1 <- data1 %>%
  filter(Health_Region!="North East" , Health_Region!="North West")

#eliminate the levels of the removed health regions
data1 <- droplevels(data1)

```

# Raking

```{r,raking}
Ages<-data.frame(Age_group_ord=c("16_34",
                             "35_54",
                             "55_and_over"),
                 Freq=c(3442815,
                        3725233,
                        4088342))

Races<-data.frame(Race=c("arab_middle_eastern",
                         "black",
                         "east_asian_pacific_islander",
                         "indigenous",
                         "latin_american",
                         "mixed",
                         "other",
                         "south_asian",
                         "white_caucasian"),
                  Freq=c(212782,
                         638346,
                         886592,
                         376558,
                         197020,
                         130033,
                         705333,
                         1166361,
                         9118079 ))

Incomes<-data.frame(Income=c("under_25000",
                             "25000_59999",
                             "60000_and_above"),
                    Freq=c(682331,
                           1395676,
                           3091164
                    ))

# corrections for Health Regions, without the North East and North West regions

Health_Regions<-data.frame(Health_Region=c(
  #"North West",
  #"North East",
  "West",
  "East",
  "Central",
  "Toronto"),
  Freq=c(#232299,
    #557000,
    4095589,
    3742520,
    5032410,
    1440644
  )
)

data1<-data1 %>%
  mutate(first_dose_m=case_when(
    first_dose=="yes"~1,
    first_dose=="no"~0
  ))

data1<- within(data1, Age_group_ord <- relevel(Age_group_ord, ref = "16_34"))
data1<- within(data1, Race <- relevel(Race, ref = "white_caucasian"))
data1<- within(data1, Health_Region <- relevel(Health_Region, ref = "Toronto"))
data1<- within(data1, Income <- relevel(Income, ref = "60000_and_above"))

a1<-svydesign(id=~1,data=data1)

## because there are no sampling weights, each observation as a probability of 1 of being sampled at this stage.

## The next line uses the survey design object a1, and the variables we will adjust for, with the population corrections
## from the data frames from above.

a1_rake<-rake(a1,
              sample=list(~Age_group_ord,
                          ~Race,
                          ~Income,
                          ~Health_Region),
              population=list(Ages,
                              Races,
                              Incomes,
                              Health_Regions))

#model

m1<-svyglm(first_dose_m~Age_group_ord+Income+Race+Health_Region,
           design=a1_rake,
           family = quasibinomial(), 
           control= list(maxit=25))

m2<-svyglm(first_dose_m~Age_group_ord+Income+Race+Health_Region+Race*Income,
           design=a1_rake,
           family = quasibinomial(), 
           control= list(maxit=25))

m3<-svyglm(first_dose_m~Age_group_ord+Income+Race+Health_Region+Race*Health_Region,
           design=a1_rake,
           family = quasibinomial(), 
           control= list(maxit=25))

m4<-svyglm(first_dose_m~Age_group_ord+Income+Race+Health_Region+Race*Income+Race*Health_Region,
           design=a1_rake,
           family = quasibinomial(), 
           control= list(maxit=25))

t1<-m1 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()

t2<-m2 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()

t3<-m3 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()

t4<-m4 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()

tm<-tbl_merge(tbls=list(t1,t2,t3,t4),tab_spanner = c("No interactions","Race x Income","Race x Health Region", "Race x Income, Race x Health Region "))

tm

t5<-data1 %>% 
  tbl_cross(row = Race, col = Income)

t6<- data1 %>%
  tbl_cross (row=Race, col= Health_Region)

tbl_stack(tbls=list(t4,t5),group_header = c("Race and Income", "Race and Health Region"))

```