### 2023-02-24 manuscript/04-discussion.qmd

# Discussion

The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022], and their availability towards the end of 2020 in certain countries made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021]. However, despite previous successful vaccination campaigns that were crucial to control diseases such as smallpox and polio [@kayser2021], vaccination efforts in the case of Covid-19 have faced multiple challenges that have complicated the achievement of global immunity.

Among the different challenges faced by Covid-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and inequity in vaccine access between low and high income countries [@tamey2022]. However, it is also well established that even in the case of high income countries that have had ample access to vaccines since 2020, such as the US, the UK, and Canada, there have been challenges in vaccination efforts due to differences in vaccine uptake among different segments of the population. More specifically, lower vaccine uptake has been associated with socio-economic factors such as race (i.e., identifying as Black, Asian, Indigenous) and household income (typically within lower income brackets) [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021]. Reasons given for this association have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020].

In addition, vaccine uptake is influenced by geography as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021]. However, in the case of Canada, most studies that have analyzed spatial differences in vaccination have been focused in country-wide or province-wide estimates [@lavoie2022;@guay2022]. Therefore, we explored spatial and socio-economic determinants of vaccination status in the province of Ontario. This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Networks (LHINs) and their incorporation into larger Health Regions [@dong2022]. The idea behind the change aimed at reducing the inequalities in healthcare that were identified under the LHIN approach [@tsasis2012], and thus, examining differences in vaccination between the Health Regions can provide decision makers with information regarding intra-provincial health disparities that may need to be addressed in the future.

Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the Covid-19 vaccine (@tbl-descriptive-stats). This is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022]. However, we identified differences in vaccination based on socio-economic and geographical factors. First, our results show significant differences in vaccination odds in individuals with a household income below CAD 60,000, and in individuals belonging to minority racial or ethnic groups. Those who identified as Black, Latin American, or  belonging to different minorities not included as categories in the survey (Southeast Asian, Filipino, West Asian, and Minority not identified elsewhere) had vaccination odds below 33% when compared to individuals that identified as White/Caucasian (@tbl-model). Geographically, significantly higher vaccination odds were identified in the West Health Region, which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.

The lack of difference in vaccination uptake across different age groups is consistent with other studies that have shown similar results [@macdonald2021;@guay2022]. In the case of Ontario, where there was around 90% of vaccination uptake in the age group above 65 years old, and xxx percentages in xxx group and xxx group, these results show an overall high vaccination uptake at the provincial level.
These results are consistent with other studies that have shown an association between lower vaccination uptake, identifying as part of a racial minority, or having a low household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].

Although there was no significant difference between vaccination odds among the Health Regions, there was a significant lower vaccination odds in the Central Health Region in individuals with a household income between CAD 15,000-24,999 (OR=0.4, @tbl-model). The Central Health Region comprises the former Central, Central West, Mississauga Halton, and North Simcoe Muskoka LHINs with 48.3% of immigrant population and 47% of the population identifying as visible minority, the highest percentage among all Health Regions [@ontariohealth]; moreover, the fact that the Central West LHIN, now within the Central Health Region, had the highest proportion of low-income families in 2018 (12.9\% of males and 16.7\% of females among the lowest 20% of Canadian family incomes [@buajitti2018])

- Influence of geography

- ontario and Health Regions: find information about poverty and racial groups.

- Results

- Limitations