### 2023-04-16 manuscript_v3/05-conclusion.qmd

# Conclusion

This study explored differences in COVID-19 vaccination across the province of Ontario between late 2021 and early 2022 by taking into consideration socio-economic factors, such as income and race, their interactions, and the Health Regions within the province. Our results show that, during the period analyzed, significant differences in vaccination existed across different visible minority groups, income brackets, and Health Regions, showing  intra-provincial disparities in vaccine uptake. As the COVID-19 continues around the world, it important that future public policies take into consideration how to adequately reach individuals within minority groups that live across geographical areas where less probabilities of being vaccinated are likely. At the moment, this is an ongoing issue that needs to be addressed to ensure a more homogeneous outcome from the pandemic.