### 2023-04-28 manuscript_v4/03-results.qmd
need to fix figure to exponentiated

# Results

## Sample Characteristics

@tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis with regard to vaccination status. A total of 3,549 observations were used for analysis, from which 27\% (958) corresponded to individuals that reported not having received the first dose of the vaccine. Respondents between 16 and 34 years of age, and those that lived within the Health Region of Toronto had the highest number of observations in the data (42.8\% and 37.3\% of the total, respectively). Moreover, individuals who identified as White or Caucasian reported the highest vaccination uptake (82\%), in contrast to respondents who identified as which had had the lowest vaccination uptake rate (60\%) across all groups. From all the covariates, significant associations were identified between income, race, and vaccination rate.

```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination Status)"
#| message: false
#| warning: false
#| echo: false

source(here::here("code","data_preparation.R"))

#create summary table

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Month,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income (CAD)")
  ) %>% add_p(test = everything () ~ "chisq.test")%>%
  bold_labels()

data_summary %>%
  modify_header(label = "**Variable**") %>%
  as_kable_extra(booktabs = TRUE,linesep="") %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down"` argument here.
  kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)

# test<-apply(xtabs(~first_dose_m + Race,clean_data), 2, chisq.test)
# 
# 
# clean_data %>%
#   filter(income_ord %in% c("60000 and above"))%>%
#   tbl_summary(percent="row",
#               by= first_dose_m,
#               missing="no",
#               include=Health_Region,
#               label=list(
#                  income_ord="Income (CAD)")
#   ) %>% add_p()%>%
#   bold_labels()

# clean_data %>%
#   filter(income_ord %in% c("25000-59999"))%>%
#   tbl_summary(percent="row",
#               by= income_ord,
#               missing="no",
#               include=(income_ord),
#               label=list(
#                  income_ord="Income (CAD)")
#   ) %>% add_p()%>%
#   bold_labels()

```

## Multivariate Regression

@fig-model-uncorr shows the estimates from the logistic regression model of vaccination status for the uncorrected data, whereas the estimated obtained from the corrected data appear in @fig-model-corr. The results show significantly higher odds of vaccination in individuals with a high household income, and those that identified as White or Caucasian when compared with individuals in lower income brackets or that identified as part of equity-deserving groups.

Specifically, significantly lower odds of vaccination were identified for those with a household income under CAD 25,000 (OR=0.374, CI=[0.251,0.559]) and those with an income between CAD 25,000 and 59,999 (OR=0.586, CI=[0.390,0.882]).  Additionally, individuals who identified as Arab/Middle Eastern, Black, or Latin American, or that belonged to "Other" racial groups, which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and CIs=[0.14,0.68], [0.17,0.60], [0.11,0.66], and [0.12,0.41], respectively). Regarding Health Regions, individuals that reported living in the West Health Region (which includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, CI=[1.04,2.29]).

Interestingly, higher odds of vaccination were also estimated in the case of individuals in equity-deserving groups with a household income below CAD 25,000 when compared to those with a household income above CAD 60,000.  This held true in the case of Arab/Middle Eastern (OR=3.08, _p_=0.013), Black (OR=3.15, _p_=0.004), and Latin American (OR=2.81, _p_=0.041) individuals, as well as respondents who belonged to "Other" minority groups  (OR=4.63, _p_<0.001). Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.96, _p_<0.001).

For the interaction of Health Region and race, significant odds of vaccination were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44, _p_=0.046), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41, _p_=0.032 and _p_=0.037, respectively).

```{r,model-uncorr-fig}
#| message: false
#| warning: false
#| echo: false
#| fig-align: left
#| fig-width: 7
#| fig-height: 7
#| label: fig-model-uncorr
#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model. Only statistically significant interaction terms are shown. Full estimates can be found in the Appendix."

#call the script for raking
source(here::here("code","raking.R"))

#logistic regression
source(here::here("code","regression.R"))

#create dataframe from modelplot, remove names of factors, drop the intercept

mplot1<-modelplot(m0,draw=FALSE,
                  coef_omit = "Intercept", exponentiate = TRUE) %>%
  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))

#add column by factor 
mplot1$factor<-c(rep("Age",2),
                 rep("Month",2),
                 rep("Income",2),
                 rep("Race/Ethnicity",8),
                 rep("Health Region",3),
                 rep("Income and\nRace/Ethnicity",16),
                 rep("Race/Ethnicity\nand\nHealth Region",24)
                 )

#order factors for plot
#mplot1$term_ord <- factor(mplot1$term,ordered=TRUE,levels=mplot1$term)

mplot1$factor <- factor(mplot1$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))

# remove non-significant interaction factors

test<-mplot1%>%
  filter(str_detect(term,':') & p.value>0.05)

#remove non-significant interactions from dataframe
mplot1<-anti_join(mplot1,test)

mplot1<-mplot1%>%
mutate(term=str_replace_all(term,':'," and "))

#make plot
mplot1 %>% 
ggplot(aes(y=term))+
  geom_vline(xintercept = 0, linetype="dashed",color="red") +
  geom_point(aes(x=estimate))+
  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
  scale_y_discrete(limits=rev)+
  xlab("Coefficient estimates and 95% confidence intervals")+
  ylab("")+
  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y",labeller = label_wrap_gen(multi_line = TRUE))+
  theme_classic()
```

```{r,model-corr-fig}
#| message: false
#| warning: false
#| echo: false
#| fig-align: left
#| fig-width: 7
#| fig-height: 7
#| label: fig-model-corr
#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model. Only statistically significant interaction terms are shown. Full estimates can be found in the Appendix."

#create dataframe from modelplot, remove names of factors, drop the intercept

mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept",exponentiate = TRUE) %>%
  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))

#add column by factor 
mplot2$factor<-c(rep("Age",2),
                 rep("Month",2),
                 rep("Income",2),
                 rep("Race/Ethnicity",8),
                 rep("Health Region",3),
                 rep("Income and\nRace/Ethnicity",16),
                 rep("Race/Ethnicity\nand\nHealth Region",24)
                 )

#order factors for plot
#mplot2$term_ord <- factor(mplot2$term,ordered=TRUE,levels=mplot2$term)

mplot2$factor <- factor(mplot2$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))

# remove non-significant interaction factors

test<-mplot2%>%
  filter(str_detect(term,':') & p.value>0.05)

mplot2<-anti_join(mplot2,test)

mplot2<-mplot2%>%
mutate(term=str_replace_all(term,':'," and "))

mplot2 %>% 
ggplot(aes(y=term))+
  geom_vline(xintercept = 5.73, linetype="dashed",color="red") +
  geom_point(aes(x=estimate))+
  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
  scale_y_discrete(limits=rev)+
  xlab("Coefficient estimates and 95% confidence intervals")+
  ylab("")+
  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y")+
  theme_classic()
```