### 2023-05-12 manuscript_v5/03-results.qmd

# Results

## Sample Characteristics

@tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis. The sample contained **6,236** observations, from which 24.8\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine. Vaccination rates ranged between 71-79\% across household income brackets, age groups, Health Regions, and the months considered in the survey. However, the highest vaccination rates in each category were reported by individuals in the highest income bracket (79\%), those between 16 and 34 years of age (77\%), individuals that lived in the East Health Region (77\%), and  during January of 2022 (78\%). Differences were higher between racial/ethnic groups, where the higher vaccination rate was reported by White/Caucasian individuals (84\%), against vaccination rates between 63-66\% reported in the case of Arab/Middle Eastern, Black, Indigenous, Latin American individuals, and those that reported belonging to "Other" racial groups, which included Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere.

```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination Status)"
#| message: false
#| warning: false
#| echo: false

source(here::here("code","data_preparation_May_08.R"))

#create summary table

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Month,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income (CAD)")
  ) %>% add_p(test = everything () ~ "chisq.test")%>%
  bold_labels() %>%
  modify_table_styling(
    columns=label,
    rows=label=="Other",
    footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
  ) %>%
  modify_header(label = "**Variable**") %>%
  as_gt()

data_summary

# test<-a1_rake$variables
# 
# a1_rake %>%
# tbl_svysummary(percent="row",
#               by= first_dose,
#               include=c(income_ord,
#                         Age_group_ord,
#                         Health_Region,
#                         Month,
#                         Race),
#               label=list(Health_Region ~ "Health Region",
#                  Age_group_ord ~ "Age Group",
#                  income_ord="Income (CAD)")
#   ) %>% add_p(test = everything () ~ "chisq.test")%>%
#   bold_labels() %>%
#   modify_table_styling(
#     columns=label,
#     rows=label=="Other",
#     footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
#   ) %>%
#   modify_header(label = "**Variable**") %>%
#   as_gt()
#

```

## Multivariate Regression

@fig-models presents the estimates (as odd ratios) from the logistic regression models for vaccination status using the socio-demographic factors collected by the survey, and their interactions. Generally speaking, lower odds of vaccination were identified in both cases in individuals characterized by a low household income, or that identified as part of equity-deserving groups. However, the magnitude of the estimates differed between the uncorrected and corrected models and more importantly, certain estimates were not deemed to be statistically-significant after the correction, in contrast to the estimates from the uncorrected model. Specifically, the corrected model showed no significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region (@fig-models,B).
<!-- \newpage -->

<!-- \begin{sidewaysfigure} -->

```{r,models}
#| message: false
#| warning: false
#| echo: false
#| fig-align: left
#| fig-width: 15
#| fig-height: 12
#| label: fig-models
#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model. Only statistically significant interaction terms are shown. Full interaction terms can be found in Supplementary Figure A-3."

#call the script for raking
source(here("code","raking.R"))

#logistic regression
source(here("code","regression.R"))

# call script that creates figure for uncorrected model
source(here("code","uncorr_model.R"))

#call script that creates figure for corrected model
source(here("code","corr_model.R"))

fig_mod<-uncorr_mod+corr_mod+plot_layout(guides = "collect")+plot_annotation(tag_levels = "A")&
    theme(plot.tag = element_text(size=18,face='bold'),
          #axis.title = element_text(size=11),
          axis.text = element_text(size=12),
          strip.text=element_text(size=12))

fig_mod

#table for comparison
#tbl_merge(tbls = list(tbl_regression(m0,exponentiate=TRUE),tbl_regression(m2,exponentiate = TRUE)))
```

<!-- \end{sidewaysfigure} -->

However, significantly lower odds of vaccination were identified in the corrected model for those with a household income under CAD 25,000 (OR=0.37, CI=[0.27,0.51]) and those with an income between CAD 25,000 and 59,999 (OR=0.58, CI=[0.42,0.81]).  Additionally, individuals who identified as Arab/Middle Eastern, Black, Latin American, of mixed background, or that belonged to other racial groups (a category that included Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere), had significantly lower odds of vaccination than those in the White/Caucasian group (ORs and CIs=0.28 [0.16,0.51], 0.27 [0.16,0.45], 0.40 [0.21,0.76], 0.53 [0.30,0.92], 0.23 [0.15,0.36]). Additionally, individuals that reported living in the Central and West Health Regions had higher odds of vaccination than those in the Health Region of Toronto (ORs and CIs=1.61 [1.10,2.34], and 1.59 [1.16,2.19], respectively).

Interestingly, individuals in equity-deserving groups with a household income below CAD 25,000 had higher odds of vaccination (when compared to those with a household income above CAD 60,000).  This held true in the case of Arab/Middle Eastern (OR=34, CI=[1.70,6.79]), Black individuals (OR=3.81, CI=[2.05, 7.09]), or those in other racial or ethnic groups (OR=3.19, CI=[2.00,5.09]). Additionally, individuals with an income between CAD 25,000 and 59,999 in the Arab/Middle Eastern and other racial ethnic groups had higher odds of vaccination (ORs and CIs=6.96 [2.67,18.16], and 3.5 [1.85,6.62]).

Finally, significantly lower odds of vaccination were identified (when compared to the Toronto Health Region) for Black individuals in the Central Health Region (OR=0.39, CI=[0.2,0.75]), Arab/Middle Eastern individuals in the East Health Region (OR=0.41 [0.17, 0.98]), and in the Indigenous and mixed groups in the West Health Region (ORs and CIs=[0.31 [0.14, 0.7] and 0.38 [0.19, 0.76], respectively).

```{r,model-corr-fig}
#| message: false
#| warning: false
#| echo: false
#| fig-align: left
#| fig-width: 7
#| fig-height: 7
#| label: fig-model-corr
#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model. Only statistically significant interaction terms are shown. Full interaction terms can be found in Supplementary Figure A-4."

#call script with code for figure
source(here("code","corr_model.R"))

#call figure
corr_mod
```