### 2023-05-16 manuscript_v5/01-background.qmd
updating discussion

# Background

As of May of 2023 there have been 765 million confirmed cases of COVID-19 around the world, including 6.8 million deaths [@WHO-Covid]. Although this disease is no longer categorized as a global health emergency by the World Health Organization (WHO) [@rigby2023], there is ongoing concern due to continued transmission, surges in cases and deaths due to new variants [@un2023], and weaknesses in health systems around the world that could be exploited by a novel virus or another public health emergency in the future [@mackey2021].

In particular, a major weakness that has received attention during the pandemic has been related to inequalities in vaccine uptake. The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic due to the start of vaccination campaigns in certain parts of the world toward the end of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020]) but inequalities in vaccine uptake made these pharmaceutical interventions ultimately unable to replicate the experience of smallpox, where vaccination on a global scale and was crucial to control this disease [@kayser2021].

This problematic is a multifaceted issue resulting from a combination of factors, among which are failed public health measures [@li2021], inequality in vaccine access between high- and low-income countries [@gerretsen2021; @tamey2022], and vaccine hesitancy [@nafilyan2021]. Furthermore, it is well established that this issue has affected in particular individuals in certain equity-deserving groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].

Reasons given for this inequality have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].  However, it is important to  also consider that vaccination uptake can be influenced by geographical (spatial) factors. In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], spatial differences in vaccine access and supply, vaccination location availability, and lack of prioritization of areas where vulnerable groups reside [@bogoch2022; @nguyen2021]. Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022].

However, such analyses have been carried mostly in territories outside of Canada, where available studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022]. Therefore, there is a need for studies that explore spatial differences in vaccination within the Canadian territory and that consequently, can help identify disparities that need to be addressed within specific areas in each province.

This need is specially important in the case of Ontario, the most populated province of Canada. Between 2007 and 2019, Ontario managed healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integration Networks (LHINs), which aimed to provide an integrated health system for the province. However, this approach was complex and bureaucratic, and resulted in excessive expenditures, disparities in mortality rates, the deterioration of certain performance indicators such as wait times and hospital readmissions, fragmented electronic health systems, the decline of performance indicators, and inequities in health services access [@tsasis2012;@muratov2018;@dong2022;@lysyk2015;@lysyk2016]. Therefore, with the intent of better organizing and delivering care in late 2019 the provincial government eliminated the LHINs and incorporated the areas covered by them into six larger Health Regions (North East, North West, Central, Toronto, West, and East) [@dong2022].

Because the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat. In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023]. Therefore, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.

Therefore, in this study we hypothesized that there were differences in vaccination uptake between the different Health Regions of Ontario between October of 2021 and January of 2022. By including socio-economic factors in our analysis, we aimed at identifying in which groups these differences were significant in order to provide an assessment of the current state of healthcare access in Ontario.