### 2023-05-16 manuscript_v5/04-discussion.qmd

# Discussion

had significantly higher odds of vaccination than those in the Health Region of Toronto

In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its missing of improving health access for the inhabitants of Ontario.

Our results show differences in odds of vaccination across Ontario in certain socio-demographic groups. Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White/Caucasian (@fig-models). These results are consistent with previous studies that have shown lower vaccination rates in individuals with the same socio-demographic characteristics [@guay2022;@muhajarine2021;@carter2022;@hussain2022].

Lower vaccine uptake in the socio-demographic groups indicated above may be influenced in part, by vaccine hesitancy and refusal, which have been associated in equity-deserving Canadian individuals to concerns on vaccine safety, effectiveness, and experiences of racial discrimination in health settings [@basta2022;@cnat2022a;@cnat2022b;@cnat2023]. However, it has been shown that structural barriers also play an important role in vaccination uptake. In the case of equity-deserving individuals, such barriers include complex scheduling systems, language barriers, lack of adequate public transportation, and lack of accessible vaccination sites [@njoku2021]. In this regard, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 in Toronto and other regions of Ontario around the time covered by the survey [@bogoch2022;@iveniuk2021], and that pharmacies in the Peel region (an area identified as a "hotspot" with high numbers of essential workers and multigenerational households) could not keep up with demand [@gill2022]. This suggests disparities in vaccine accessibility that affected in particular equity-deserving individuals in Ontario at the time of the survey. However, because to the best of our knowledge there seems to be a very limited amount of literature on this topic in the context of Ontario, there is an ongoing need of future studies that examine the longitudinal impact of vaccine accessibility and structural barriers that affect equity-deserving groups within the province.

Interestingly, whereas overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in lower income brackets when compared to the CAD 60,000 and over bracket (Supplementary Figure A-5). Specifically, individuals in low income brackets that belonged to Arab/Middle Eastern, Black, or other minority groups had higher odds of vaccination that their peers with an income above 60,000 CAD.

This result is likely reflects in part the fact that individuals in racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include grocery store, gas station, warehouse, distribution, and manufacturing workers, all being occupations for which an income within the significant brackets identified in the analysis is to be expected. In Ontario, these workers had priority for COVID-19 vaccination [@mishra2021]; and there is evidence of interventions by vaccination staff in certain parts of the province to encourage vaccination uptake by these individuals [@gill2022]. These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation from individuals in equity-deserving groups played an important role in increasing the odds of vaccination in the province.

However, the results also indicate that the place of habitation within the province for certain equity-deserving groups also affected the odds of vaccination (interaction terms of Health Region and Race, @fig-models). Specifically, this held true in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models]. For these individuals, vaccination odds were lower when compared to the Toronto Health Region (Supplementary Figure A-6). Although it is likely that these findings have multiple causes, we indicate next some contributing factors that might help provide context in each case.

First, in the case of Indigenous, mixed, and Arab/Middle Eastern individuals is useful to disaggregate the data by LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses. Interestingly, for the first two groups most of the observations in the survey (for West Health Region) came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs. Previous research has identified health disparities in these (mostly rural) regions, such as unequal distribution of primary care providers, increased mortality, and low pharmacist availability [@shah2019;@crighton2015;@timony2022].

Furthermore, there is an ongoing challenge for the health system of the province with regard to personalized healthcare for marginalized individuals. For example, the West Health Region has only two Aboriginal Health Access Centres in contrast to an Indigenous population of about 100,000 individuals [@ontariohealth]. Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada where these rationales have been associated to lower vaccination among these groups [@smylie2022;@eissa2021]. Taken together, these aspects indicate ongoing multiple healthcare disparities that continue to affect equity-deserving individuals across the province.

The West Health Region is located in the leftmost southern part of the province, including the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent. The Central Health Region includes the regions of Peel and Halton, the regional municipality of York, the district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of Etobicoke.

There are some limitations to the present study. First, the data collection design, which allowed respondents to withdraw from the survey at any point, resulted in a high number of unique entries in the survey with multiple missing answers. Because we focused on entries that had complete observations in the covariates of interest for our analysis, it is possible that some information was not considered by excluding observations that had information in other variables (such as work from home, or number of persons in the household). However, we attempted to minimize this possibility by correcting the dataset using information from the Census. More granular corrections, which for example could be based on demographic information by municipality, could be used in the future to obtain a more accurate approximation to the population totals of the province. Moreover, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map). Although low representation from these areas is based on the fact that these regions only account for 5\% of the total population of Ontario, these regions are the home to more than 100,000 individuals that identify as Indigenous [@ontariohealth], a minority group that has historically suffered from reduced access to health care and discrimination [@mosby2021]. Therefore, there is a need for additional studies that focus on these low-populated Health Regions in Ontario where disparities in vaccination might be significant and understudied.

The results in this study are based on self-reported data, where bias might be present. However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid assessment of vaccination in the province. Finally, this study focused on first-dose vaccination status  within a relatively short time window, and therefore can only provide a snapshot of the societal dynamics behind the pandemic. Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.