### 2023-05-17 manuscript_v5/04-discussion.qmd
discussion updated

# Discussion

In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its mission of improving health access for the inhabitants of Ontario.

Our results show that indeed, there were differences in vaccination odds across Ontario in certain socio-demographic groups. Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White/Caucasian (@fig-models). These results are consistent with previous studies that have shown lower vaccination rates in individuals with the same socio-demographic characteristics [@guay2022;@muhajarine2021;@carter2022;@hussain2022].

Lower vaccine uptake in the socio-demographic groups indicated above may be influenced in part, by vaccine hesitancy and refusal, which have been associated in equity-deserving Canadian individuals to concerns on vaccine safety, effectiveness, and experiences of racial discrimination in health settings [@basta2022;@cnat2022a;@cnat2022b;@cnat2023]. However, it has been shown that structural barriers also play an important role in vaccination uptake. In the case of equity-deserving individuals, such barriers include complex scheduling systems, language barriers, lack of adequate public transportation, and lack of accessible vaccination sites [@njoku2021]. In this regard, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 in Toronto and other regions of Ontario around the time covered by the survey [@bogoch2022;@iveniuk2021], and that pharmacies in the Peel region (an area identified as a "hotspot" with high numbers of essential workers and multigenerational households) could not keep up with demand [@gill2022]. This suggests disparities in vaccine accessibility that affected in particular equity-deserving individuals in Ontario at the time of the survey. However, because to the best of our knowledge there seems to be a very limited amount of literature on this topic in the context of Ontario, there is an ongoing need of future studies that examine the longitudinal impact of vaccine accessibility and structural barriers that affect equity-deserving groups within the province.

Interestingly, whereas overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in lower income brackets when compared to the CAD 60,000 and over bracket (Supplementary Figure A-5). Specifically, individuals in low income brackets that belonged to Arab/Middle Eastern, Black, or other minority groups had higher odds of vaccination that their peers with an income above 60,000 CAD.

This result is likely reflects in part the fact that individuals in racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include grocery store, gas station, warehouse, distribution, and manufacturing workers, all being occupations for which an income within the significant brackets identified in the analysis is to be expected. In Ontario, these workers had priority for COVID-19 vaccination [@mishra2021]; and there is evidence of interventions by vaccination staff in certain parts of the province to encourage vaccination uptake by these individuals [@gill2022]. These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation from individuals in equity-deserving groups played an important role in increasing the odds of vaccination in the province.

However, the results also indicate that the place of habitation affected the odds of vaccination for certain equity-deserving groups (interaction term of Health Region and Race, @fig-models). Specifically, this held true in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models]. For these individuals, vaccination odds were lower when compared to the Toronto Health Region (Supplementary Figure A-6). We indicate next some contributing factors that might help provide context in each case.

First, it is useful to analyze the data using the LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses. Interestingly, for Indigenous and mixed individuals most of the observations in the survey for West Health Region came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs. Previous research has identified health disparities in these (mostly rural) regions, such as unequal distribution of primary care providers, increased mortality, and low pharmacist availability [@shah2019;@crighton2015;@timony2022].

Furthermore, there is an ongoing challenge for the health system of the province with regard to personalized healthcare for marginalized individuals. For example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated 100,000 Indigenous individuals living in the area [@ontariohealth]. Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada. Indeed, these rationales have been associated to observed lower vaccination rates among these groups [@smylie2022;@eissa2021]. Taken together, this suggests that healthcare disparities that are specific to certain socio-demographic groups are associated with lower vaccination uptake.

There are some limitations to the present study. First, the data collection design, which allowed respondents to withdraw from the survey at any point, and that deployed the questions in a random manner resulted in an elevated number of missing observations without a definite pattern and complicated the implementation of sensitivity analyses. Therefore, we focused on entries that had complete answers, and corrected the data using population-wide information from the Census. However, more granular corrections would be needed to obtain more accurate estimates. For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates that other racial groups. Correcting for each racial/ethnic group in each Health Region can provide a more accurate estimation of region-wide vaccination rates. Unfortunately, at the moment this correction cannot be implemented as such stratification is not currently available in the Census data.

Additionally, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map). Low representation is expected as these regions only account for 5\% of the total population of Ontario, but in contrast, they have the highest proportion of Indigenous inhabitants [@ontariohealth]. In the context of personalized care, there is a need for collecting data that focuses on these Health Regions where additional health disparities might be present and possibly understudied.

The results in this study are based on self-reported data, where bias might be present. However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid assessment of vaccination in the province. Finally, although higher vaccination odds identified for January of 2022 in the model are in accordance with province-wide trends reported by Public Health Ontario (which show a 4\% increase between early December and January, in contrast to a 2.5\% increase between October and November [@ontario-covid]), these results are only a snapshot when compared to the overall duration of the pandemic.

Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.