### 2023-05-18 manuscript_v5/05-conclusion.qmd

# Conclusion

The implementation of the Health Region model in Ontario aimed at reducing the bureaucratic complexity and disparities in health identified under the LHINs. However, the complex socio-demographic landscape of the province poses multiple challenges to the aim of the Health Region model of improving healthcare for all Ontarians. In particular, because the Health Regions cover larger geographical areas with various degrees of rurality and representation of equity-deserving groups, there is a need for decision-makers to consider that the evidence collected during the COVID-19 indicates a lack of infrastructure that can support personalized care for marginalized individuals

This study explored differences in COVID-19 vaccination across the province of Ontario during the last quarter of 2021 taking into consideration socio-economic factors, such as income and race, their interactions, and the Health Regions within the province. Our results show that during the period analyzed, differences in vaccination uptake existed across multiple equity-deserving groups in the province, and that these differences were significant in two of the Health Regions analyzed. It is important that future public health policies in Ontario take into consideration how to adequately reach individuals from equity-deserving groups that might live in areas of the province where access to healthcare might be difficult. Only in this way the goal of the Health Region model, which aims at reducing disparities, will become successful.