### 2023-07-12 manuscript_v6/02-methods.qmd

# Methods

## Data and Methods {#sec-data}

We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto (under Research Ethics Board approval #40999), and which ran between September 30th, 2021 and January 17th,2022.

Briefly, the survey was deployed using Random Domain Intercept Technology (RDIT), a methodology for internet surveys developed by the company commissioned to run the survey (RIWI Corp., Toronto, Canada), and that has been used in the area of public health research to examine trends in vaccination rates[@sargent2022], ratings of care quality[@kim2021], and perceptions of vaccine efficacy[@seeman2010]. In the case of the survey used for this study, internet users whose device meta-data indicated their presence in the province of Ontario had a random chance of being redirected to the survey after they had clicked on a registered but commercially inactive web link, or after they had typed in a web address for a site that was dormant but that was temporarily managed by RIWI. Users then decided whether to anonymously participate in the survey, and those that participated were able to exit the survey at any time. After the survey closed, regardless if it was complete or incomplete, access was denied to any further users with the same internet protocol address (IP), effectively allowing each user only one opportunity to participate in the survey. Users who indicated they were under the age of 16 were exited from the survey without creating a record. Finally, the personal identifier information from each user that participated in the survey was automatically scrubbed and replaced by a unique ID.

Survey users entered their socio-economic information (age, income, and racial/ethnic group), and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no" (@tbl-descriptive-stats). Of notice, because the racial/ethnic categories of the survey did not match exactly the categories from the 2016 Census, we used a combination of sources to re-group certain categories in order to obtain estimates that could be used to correct the data. Further details can be found in the Appendix.

Additionally, the survey automatically collected the geographical location of the respondent (using the nearest municipality,  as shown in @fig-map), and the date of access to the survey. The original dataset contained 39,029 observations. However, the number of complete observations was much lower that the total number of observations due to the survey design, which allowed respondents to exit at any time and deployed the questions randomly. We selected the observations with complete answers (6,343 observations, or 16.25\% of the total) for our analysis. It should be noted that this response rate is similar to response rates observed in previous studies that have used the same type of survey instrument and that have ranged between 15\% and 22\%[@seeman2010;@kim2021]. Next, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (254 observations corresponding to the North West and North East Health Regions). After all the preliminary analyses, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering between October 1st, 2021 and January 17, 2022. The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.

```{r,map}
#| label: fig-map
#| fig-cap: "Geographic representation of the data collected by the _Survey of COVID-19 related Behaviours and Attitudes_, collected by the Fields Institute in Ontario. Municipalities from where survey participants provided answers appear as points, color indicates number of observation obtained from each city. The Health six Regions are color-coded and labelled sequentially. Internal boundaries within certain Health Regions indicate areas previously covered by the Local Integrated Health Networks (LHINs)."
#| message: false
#| echo: false
#| out-height: 100%
#| out-width: 100%

knitr::include_graphics(here("data","map_data","map_June_01.pdf"))

```

## Statistical analyses

We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a]. Because we identified differences in representativity between the survey data and the estimates from the 2016 Census for the variables considered in the analysis, and used used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data. Details regarding the correction can be found in the Appendix. All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], `quarto` [@quarto], `modelsummary` [@modelsummary], and `gtsummary` [@gtsummary].