### 2023-07-28 manuscript_v7/05-conclusion.qmd

# Conclusion

The evidence collected in this study shows differences in the uptake of the first dose of the COVID-19 vaccine in Ontario between October 2021 and January 2022 in underrepresented groups, which had significantly lower odds of vaccination when compared to White/Caucasian individuals. However, although overall vaccination uptake was lower in underrepresented individuals in the province as a whole, the odds of vaccination differed within certain income levels and Health Regions in these groups. Specifically, those that reported a low household income had higher vaccination odds when compared to individuals in the same racial/ethnic group in a higher income bracket. These results highlight the complex landscape of the province, where varying degrees of rurality exist in conjunction with a socio-demographic makeup that is unique to each of the Health Regions.

Personalized care is an area that could be further developed to improve vaccination uptake in underrepresented individuals in the future. As we have shown in the discussion, there are currently a limited number of centers focused on community-led healthcare in the province. Our results showed the existence of differences in vaccination uptake between certain Health Regions and therefore, improving personalized care for underrepresented individuals can serve as a point to improve trust and facilitate vaccine access in these marginalized communities.

Overall, there are ongoing challenges for healthcare access across different segments of the population of Ontario. Future studies can be focused at more in-depth analyses of vaccination uptake within the different socio-demographic groups in each Health Region, in order to provide decision-makers with information that can serve to carefully consider and address how to improve vaccine access to marginalized communities in Ontario in the event of a future public health emergency.